A Comparison of Parallel Graph Processing Implementations

### 2017-08-14 papers/publication/paper-pollard.tex
This is the vesion I turned into Cluster'17

\subsection{Scalability}
Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong efficiency of the BFS algorithm in different packages on a Kronecker graph of scale $23$. The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads. Because of timing considerations, only four trials were run for these experiments.
\begin{figure}[htb]
	\centering
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup23.pdf}
		\caption{Speedup of BFS for a scale-20 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
	\label{fig:bfs-speedup}
\end{figure}

@@

Comparison of these two datasets in \figurename~\ref{fig:epg-realworld} brings up some surprising variations in the data. For example, PowerGraph is faster for SSSP. This could be because of the efficient edge-cut partitioning scheme in place on PowerGraph which can more efficiently deal with the high degree vertices present on the denser Dota-League graph. However, this comes with a significant overhead; PowerGraph is slower for SSSP than the other platforms. GraphBIG seems to have the widest variation; it is by far the slowest for PageRank but is the fastest for the Dota-League BFS, beating out even GAP which uses a more modern algorithm for BFS called $\Delta$-stepping \cite{Meyer:2003:delta}. One reason for this lack of performance from GAP is our lack of tuning; we use the default $\Delta = 1$ which is known not to be optimal. One possible explanation for the poor performance of GraphMat compared is the overhead of the sparse matrix operations. These may pay off for larger datasets and we see good performance on the denser Dota-League dataset for GraphMat across all algorithms.

@@

Recent advances in parallel SSSP and BFS algorithms make use of $\Delta$-stepping. This consists of the algorithm being parametrized by some $\Delta > 0$ for which requires tuning for a given architecture and graph structure to achieve optimal performance~\cite{Panitanarak:2014:SSSPPerf}. This is provided in the GAP benchmark suite as well as the Parallel Boost Graph Library and we plan to add some level of parameter tuning to the next iteration of our framework.

>>>>

\subsection{Scalability}
Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong efficiency of the BFS algorithm in different packages on a Kronecker graph of scale $23$. The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads. Because of timing considerations, only four trials were run for these experiments.
\begin{figure}[htb]
	\centering
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup23.pdf}
		\caption{Speedup of BFS for a scale-23 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
	\label{fig:bfs-speedup}
\end{figure}

@@

Comparison of these two datasets in \figurename~\ref{fig:epg-realworld} brings up some surprising variations in the data. For example, PowerGraph is faster for SSSP. This could be because of the efficient edge-cut partitioning scheme in place on PowerGraph which can more efficiently deal with the high degree vertices present on the denser Dota-League graph. However, this comes with a significant overhead; PowerGraph is slower for SSSP than the other platforms. GraphBIG seems to have the widest variation; it is by far the slowest for PageRank but is the fastest for the Dota-League BFS, beating out even GAP which uses a more modern algorithm for BFS called Direction-optiming BFS \cite{Beamer:2012:DOBFS}. One reason for this lack of performance from GAP is our lack of tuning; we use the default parameterization of $\alpha = 15$ and $\beta = 18$, which may not be optimal for all graphs. One possible explanation for the poor performance of GraphMat compared is the overhead of the sparse matrix operations. These may pay off for larger datasets and we see good performance on the denser Dota-League dataset for GraphMat across all algorithms.

@@

Advances in parallel SSSP and BFS contain parameterizations ($\Delta$ for SSSP and $\alpha$ and $\beta$ for BFS) which affects performance depending on graph structure. These are provided in GAP. We plan to add some level of heuristic parameter tuning as performed in \cite{Beamer:2012:DOBFS} to the next iteration of our framework to take advantage of these algorithmic advances.
